---
title: "Philly_Parking"
author: "Bahman Ajdari"
date: "12/3/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading libraries in to R
```{r}
library(tidyverse)
library(lubridate)
library(ggmap)
library(sf)
library(osmdata)
library(viridis)
library(beepr)
register_google("google_api_key")
```
#Loading data in to R and some cleanings
This dataset comes from tidytuesday weekly project. 
```{r}
tickets <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-12-03/tickets.csv")
beep()
tickets$violation_desc <- str_remove(tickets$violation_desc, pattern = " CC")
set.seed(2019)
sampled_tickets <- tickets %>%
  mutate(hour = hour(issue_datetime)) %>%
  mutate(month = month(issue_datetime, label = TRUE)) %>%
  sample_n(100000)

write_csv(sampled_tickets, "sample_ticket.csv")



```
#exploring dataset
Most of the fines are less than 60$ and most of them are issued by PPA and police department.
```{r}
tickets %>%
  summary()

tickets %>%
  count(violation_desc, sort = TRUE)

tickets %>%
  ggplot(aes(fine))+
  geom_histogram(binwidth = 10)+
  coord_cartesian(xlim = c(0,250))

tickets %>%
  count(issuing_agency, sort = TRUE)



```
#ANOVA test on days
The anova test shows that there is a significant difference between fine means in different days. It seems Monday and Wednesdays are more simililar regarding to fine amount. 
```{r}
tickets_processed<- tickets %>%
  mutate(hour = hour(issue_datetime)) %>%
  mutate(month = month(issue_datetime, label = TRUE)) %>%
  mutate(day=wday(issue_datetime, label = TRUE))%>%
  filter(fine<250)%>%
  drop_na(fine)

tickets_processed %>%
  filter(fine<250)%>%
  ggplot(aes(day, fine))+
  geom_boxplot()

res.aov<- aov(fine~day, data = tickets_processed)
summary(res.aov)
TukeyHSD(res.aov)

```

# loading Philadelphia map in to R
```{r}
library(maps)
phi_map <- get_map(getbb("Philadelphia"), zoom = 11)
ggmap(phi_map_stamen)
phi_map_stamen <- get_map(getbb("Philadelphia"),source = "stamen" , maptype = "terrain-lines", zoom = 11)
map_data_phi<- map_data("state")
```

#Top 5 violations on the map
```{r}
top_5<- tickets %>%
  count(violation_desc, sort = TRUE) %>%
  top_n(5)
top_5 %>%
  mutate(violation_desc=fct_reorder(violation_desc, n))%>%
  ggplot(aes(violation_desc, n))+
  geom_col()+
  coord_flip()

tickets_filtered<- tickets %>%
  filter(violation_desc %in% top_5$violation_desc) %>%
  drop_na(lat)%>%
  drop_na(lon)
ggmap(phi_map)+
  geom_bin2d(data = tickets_filtered, aes(lon, lat, fill=violation_desc), alpha=0.5)+
  scale_fill_viridis_d(option="inferno")+
  theme_void()
ggmap(phi_map_stamen)+
  geom_point(data = tickets_filtered, aes(lon, lat, color=violation_desc), alpha=0.5)
```
# What is the chance of being cited by the time of the day
it seems mid day has the more probability for being cited. Mid night has the lowest.
```{r}
time_window <-tickets %>%
  mutate(hour=hour(issue_datetime),
         wday=wday(issue_datetime, label = TRUE))%>%
  mutate(time_window=case_when(
    between(hour, 7,10) ~ "Morning",
    between(hour, 11,15) ~ "Midday",
    between(hour, 16,18) ~ "Evening",
    between(hour, 19,23) ~ "night",
    TRUE ~ "Mid Night"
  )) %>%
  filter(violation_desc %in% top_5$violation_desc)%>%
  group_by(time_window, wday) %>%
  summarise(count=n())%>%
  ungroup()

time_window %>%
  mutate(time_window=fct_reorder(time_window,count, .desc = FALSE))%>%
  ggplot(aes(time_window, wday, fill=count)) +
      geom_tile(na.rm = TRUE)
```

#Sample data
Just for the memory issues from now on I will use samle data that I made from the original dataset

#Projectind data on the map using openstreet map API

```{r}
coord <- getbb("Philadelphia")
sampled_tickets <- filter(sampled_tickets, between(lon, coord[1, 1], coord[1, 2]), between(lat, coord[2, 1], coord[2, 2]))
streets1 <- opq(coord) %>% 
    add_osm_feature('highway', c('motorway', 'primary', 'secondary', 'tertiary')) %>% 
    osmdata_sf()
beep()

streets2 <- opq(coord) %>% 
    add_osm_feature('highway', c('residential', 'living_street', 'unclassified', 'service', 'foodway')) %>% 
    osmdata_sf()
beep()
sample_ticket
ggplot(sample_ticket)+
  geom_sf(data = streets2$osm_lines, col="grey40", size=.1)+
  geom_sf(data = streets2$osm_lines, col="grey40", size=.4)+
  geom_bin2d(aes(lon, lat, fill=..count..),binwidth = 0.01)+
  geom_sf(data = streets2$osm_lines, size = 0.1)+
  scale_fill_viridis(option = "plasma")
plot(streets2)
```

