---
title: "Shelters in Toronto"
date: 2020-12-01
output: html_output
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(lubridate)
library(scales)
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2020-12-01")
shelter <- tt$shelters
```


# Weather data
Importing climate daily stats from "https://toronto.weatherstats.ca/download.html"

```{r}
library(readr)
weather <- read_csv("~/Downloads/weatherstats_toronto_daily (1).csv")%>%
  select(date, max_temperature, min_temperature, avg_temperature, precipitation)

```

# Number of shelters per organization
```{r}
shelter %>%
  count(organization_name, sort = TRUE) %>%
  mutate(organization_name=fct_reorder(organization_name, n)) %>%
  head(16) %>%
  ggplot(aes(organization_name, n))+
  geom_col()+
  coord_flip()
```

# Joining data frames
```{r}
# Making shelter df to smaller one
shelter_small <- shelter %>%
  filter(!is.na(occupancy)) %>%
  filter(!is.na(capacity)) %>%
  filter(capacity!=0) %>%
  mutate(rate= occupancy/capacity) %>%
  mutate(date= ymd(occupancy_date)) %>%
  select(id,
         date,
         occupancy,
         capacity,
         rate)
  
joined_shelter<- shelter_small %>%
  left_join(weather, by= "date") %>%
  mutate(percipitation_deg= case_when(
    precipitation== 0 ~ "None",
    between(precipitation, 0.01,10) ~ "Low",
    between(precipitation, 10.01,20) ~ "Medium",
    between(precipitation, 20.01, 40) ~ "High",
    TRUE ~ "Very High"
  ))

```

# Daily occupancy rate and tempreture
I don't see any relationship between temperature and occupancy rate 

```{r}
joined_shelter %>%
  group_by(day=day(date), month=month(date), year=year(date), percipitation_deg) %>%
  summarise(
    avg_rate=mean(rate, na.rm = TRUE),
    avg_temp= mean(avg_temperature, na.rm = TRUE),
    avg_percipitation= mean(precipitation, na.rm = TRUE)
  ) %>% 
  ggplot(aes(avg_temp, avg_rate))+
  geom_point(alpha=0.7)+
  scale_y_continuous(labels = percent_format())+
  geom_smooth(se=FALSE)+
  facet_wrap(~year)+
  labs(
    title = "Daily Occupancy Rate in Toronto's Shelter System",
    subtitle = "Based on Daily Average Temperature(Celsius)",
    y= "Average Occupancy Rate",
    x= "Average Temperature (Celsius)"
  )
```

# monthy occupancy rate

```{r}
joined_weather %>%
  mutate(month=month(date, abbr = TRUE),
         year= year(date)) %>%
  group_by(year, month) %>%
  summarise(
    n=n(),
    all_occup= sum(occupancy),
    all_capacity= sum(capacity, na.rm = TRUE),
    avg_percipitation= mean(precipitation, na.rm = TRUE),
    avg_temp= mean(avg_temperature)
  ) %>% 
  mutate (rate= all_occup/all_capacity) %>%
  ggplot(aes(as.factor(month), rate, fill=factor(year)))+
  geom_col(position = "fill")
  scale_x_discrete(label=month.name)

```

# Save Image
Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "daily.png",
  device = "png")

```
